When are Fuzzy Extractors Possible?

### 2014-10-01 trunk/info-bounds/FuzzyExtPossibility.tex
working on introduction

\title{When are Fuzzy Extractors Possible?}
\blind{
\author{Benjamin Fuller\footnote{Boston University and MIT Lincoln Laboratory. 
%The Lincoln Laboratory portion of this work was sponsored by the Department of the Air Force under Air Force Contract \#FA8721-05-C-0002.  Opinions, interpretations, conclusions and recommendations are those of the author and are not necessarily endorsed by the United States Government.  
Email: {\tt bfuller@cs.bu.edu}.}~~~~~~~~~~Leonid Reyzin\footnote{Boston University.  Email: {\tt reyzin@cs.bu.edu}.}~~~~~~~~~~Adam Smith\footnote{Pennsylvania State University. Work done in part while the author was at Boston University and Harvard University.  Email: {\tt asmith@psu.edu}.  }}
}
\begin{document}
\maketitle

\begin{abstract}
We characterize when deriving a stable key from a noisy source is possible.  We use the notation of fuzzy extractors introduced by Dodis et al. (Eurocrypt 2004).  The goal is to produce a strong $key$ from a $w$ drawn from a \emph{strong} distribution $W$ and to produce the same when given a nearby $w'$.  A fuzzy extractor is split into two algorithms $\gen$ which produces $key$ along with some helper information $p$ and $\rep$ which takes $w'$ and $p$ to reproduce $key$.  The key must remain strong in the presence of $p$.  The goal is to produce the correct key whenever $\dis(w, w')\le t$.
\anote{there is a big gap between theory and the needs of practice}

We identify a key quantity of a distribution called \emph{fuzzy min-entropy}.  Fuzzy min-entropy measures the maximum probability mass of a distribution that lies within a ball of radius $t$ in the metric space.  Ideally, key derivation should be possible whenever a distribution has fuzzy min-entropy.  In the interactive setting, fuzzy min-entropy is both a necessary and sufficient condition for security.

Current techniques for constructing fuzzy extractors lie far from this necessary condition.  We investigate whether this is inherent for both fuzzy extractors and secure sketches~(a one round information-reconciliation component).  Our constructions and impossibility results are information-theoretic.

We consider two types of fuzzy extractors and secure sketches.  The first we call \emph{distribution-aware}, the algorithms know the precise distribution they are trying to correct.  The second is \emph{distribution-oblivious} where $(\gen, \rep)$ must work for a family of distributions.

\textbf{Distribution-Aware} Whenever the fuzzy min-entropy is super-logarithmic, there exists a distribution-aware secure sketch and fuzzy extractor with super-logarithmic output entropy.  That is, fuzzy min-entropy is both a necessary and \emph{sufficient} condition for security in this setting.

\textbf{Distribution-Oblivious} The distribution-oblivious situation is more complicated.  Indeed, there is a significant difference between secure sketches and fuzzy extractors.  We show that there exist families of distributions where each distribution has fuzzy min-entropy such that any secure sketch is insecure.  However, it is possible to build a fuzzy extractor for these same family of distributions.
\bnote{come back to this.}
\end{abstract}

\section{Introduction}
\anote{need to mention early work on trying to do distribution specific. For example, Juels and Wattenberg, Tuyls et al.}
\anote{should our results be seen as justifying the distribution specific approach?}
\anote{make the point that narrowing the family $\mathcal{W}$ makes the security guarantees much more brittle}
Usually, authentication requires some high quality secret.  Traditionally, this secret is assumed to have high min-entropy.  However, many sources with sufficient entropy for authentication present an additional problem: noise.  That is, when the physical instantiation of the source is read multiple times, readings are close~(according to some metric) but not identical.  To use such sources in an authentication scheme it is often necessary to remove this noise and derive the same key from the initial and subsequent readings.

@@

A fuzzy extractor performs both these tasks non interactively~\cite{DBLP:journals/siamcomp/DodisORS08}.  Consider a particular high entropy distribution $W$.  Our goal is to derive stable keys from $W$ whenever the two readings~(denoted $w$ and $w'$ respectively) are within distance $t$.  
A fuzzy extractor consists of two algorithms: generate~($\gen$) which takes the initial reading $w$ and produces $key$ along with a public helper value $p$.  The reproduce~($\rep$) algorithm takes the subsequent reading $w'$ along with the helper value $p$ to reproduce $key$.  The $key$ should hold even if an adversary has observed $p$~(the problem is trivial if $p$ is private).  The correctness guarantee is that the key is reproduced if the distance between $w$ and $w'$ is at most $t$.

Traditionally, fuzzy extractors are constructed by using a separate information-reconciliation component, known as a secure sketch, that maps $w'$ back to $w$ and a privacy-amplification component, a strong randomness extractor~\cite{nisan1993randomness}, that maps $w$ to a random key.  A secure sketch consists of a pair of algorithms $\sketch$ takes $w$ and produces a public value $ss$, $\rec$ takes $ss$ and a nearby value $w'$ and outputs the original reading $w$.  The security losses incurred in information-reconciliation~(performed in secure sketches and fuzzy extractors) are larger than losses in privacy amplification.\bnote{should say something here about our understanding of privacy amplification and that extractors are necessary and sufficient} In the following discussion we describe limitations on fuzzy extractors.  This discussion applies to secure sketches as well. \bnote{should say you don't have to build secure sketches}

Fuzzy extractors are designed with limited knowledge of the distributions that will later be used.  Traditionally, the distribution $W$ is assumed to be an arbitrary distribution with min-entropy.  Making minimal assumptions about the distribution $W$ is desirable for two reasons: 1) for high entropy distributions, only a small part of the distribution has been observed and the remainder of the distribution is unknown 2) the adversary may have some side information about the distribution $W$ and thus the adversaries view of the distribution is different than that of the algorithm designer.  Usually fuzzy extractors support a family of distributions $\mathcal{W}$ and hope that the target distribution $W\in\mathcal{W}$.

@@

Traditional fuzzy extractors and secure sketches operate in the worst case and incur entropy loss of the size of the ball to be error-corrected.  This loss is usually linear in the input length of the string.  While, this is optimal for the uniform distribution, it is far from optimal for distributions that arise in practice.  As an example, many examples that arise in practice there are more error patterns than starting entropy.  Traditional fuzzy extractors provide no guarantee for sources of this type.  The work of Canetti et al.~\cite{canetti2014key} constructs fuzzy  extractors for some distributions with more errors than entropy.  This leads to the following natural question: \emph{when is key derivation from noisy sources possible?}

\paragraph{The price of distributional uncertainty}
As described above, fuzzy extractors are usually built for a family of distributions $\mathcal{W}$.  This strengthens the security guarantees: security holds as long as $W\in\mathcal{W}$.  In this work, we show that this uncertainty comes at a significant cost.  We consider two settings, first where the algorithms $\gen, \rep$ know the entire probability mass function of the distribution $W$.  We call this setting \emph{distribution aware}.\bnote{prior works on this setting}  The more traditional setting where security holds for a family of distributions $\mathcal{W}$ we call \emph{distribution-oblivious}.

@@

\paragraph{Positive Results} We first consider the \emph{distribution-aware} setting.  We show for a distribution-aware secure sketch for every source with super-logarithmic fuzzy min-entropy with super-logarithmic output entropy~(\thref{cons:leveling}).  Using standard techniques we can convert this secure sketch to a fuzzy extractor with super-logarithmic key output length.\footnote{Assuming exponential-hard pseudorandom generators, a super-logarithmic length key can be expanded to an arbitrary length.  In this work, we consider this minimal goal of a super-logarithmic length key.  If we require a longer output length this increases the requirement on fuzzy min-entropy.}  Thus, in the distribution-aware setting, fuzzy min-entropy is both a necessary and sufficient condition for security.

Our constructions are primarily feasibility results and do not run in polynomial time.  In the distribution-aware setting this seems necessary as most distributions have exponential description size.  The power of the distribution-aware setting is that the algorithms can access this description.  It seems unlikely that algorithms with access to a polynomial size part of the distribution description can replicate these results.

\paragraph{Negative Results}
We then show the situation is worse in the distribution-oblivious setting.  We show that distributional uncertainty is a real drawback to providing security.  We show a family of distributions~(each with super-logarithmic fuzzy min-entropy) such that any secure sketch for the family the remaining entropy is at most $1$~(\thref{thm:imposs sketch}).  This result is quite strong, it requires the secure sketch to correct only a constant number of errors and allows the secure sketch to be wrong almost a quarter of the time.

Using a result of Fuller et al.~\cite{fuller2013computational}, this negative result extends to versions of secure sketches ensure $w$ has  pseudoentropy~(\corref{cor:extension to fuzz ext}).  Under very strong assumptions, computational secure sketches that provide unpredictability of $w$ are constructible from semantically secure multi-linear encodings~\cite{BitanskyCKP14}.  Unpredictability is weaker than pseudoentropy and converting unpredictability to a key requires a special type of extractor known as a reconstructive randomness extractor~\cite{barak-computational, DBLP:conf/eurocrypt/HsiaoLR07}.

@@

Computational fuzzy extractors offer more hope for avoiding the impossibility result.  Computational fuzzy extractors for all distributions are implied by semantically secure multi-linear encodings~\cite{BitanskyCKP14}.  Furthermore, there are computational fuzzy extractors secure for the specific family of distributions used in the counter-example under weaker assumptions~\cite[Construction 5.3]{canetti2014key}.
%The situation is significantly more complicated for distribution-oblivious algorithms.  Indeed, we show a natural families to distributions where every distribution has fuzzy min-entropy such that no secure sketch can provide meaningful security.  However, it is possible to build a meaningful fuzzy extractor for this same class of distributions.

@@

\subsection{Fuzzy min-entropy: necessary for security}
\label{sec:minimal conditions}

>>>>

\title{When are Fuzzy Extractors Possible?}
\blind{
\author{Benjamin Fuller\footnote{Boston University and MIT Lincoln Laboratory. 
The Lincoln Laboratory portion of this work was sponsored by the Department of the Air Force under Air Force Contract \#FA8721-05-C-0002.  Opinions, interpretations, conclusions and recommendations are those of the author and are not necessarily endorsed by the United States Government.  
Email: {\tt bfuller@cs.bu.edu}.}~~~~~~~~~~Leonid Reyzin\footnote{Boston University.  Email: {\tt reyzin@cs.bu.edu}.}~~~~~~~~~~Adam Smith\footnote{Pennsylvania State University. Work done in part while the author was at Boston University and Harvard University.  Email: {\tt asmith@psu.edu}.  }}
}
\begin{document}
\maketitle

\begin{abstract}
Fuzzy extractors (Dodis et al., Eurocrypt 2004) convert repeated noisy readings of a high-entropy secret into the same uniformly distributed key. To help eliminate noise, they produce a non secret helper string. This helper value allows the key to be regenerated from a nearby point.  Known constructions of fuzzy extractors do not provide security for most noisy distributions that occur in practice.

The nonsecret helper string allows an adversary to learn the generated key if they can guess a point close to the original reading.  A necessary condition to derive keys from a distribution is that no adversary can guess nearby points.  We call this condition \emph{fuzzy min-entropy}.  In the interactive setting, fuzzy min-entropy is both a necessary and sufficient condition for security.  In this work, we ask whether fuzzy min-entropy is a sufficient condition for security in the fuzzy extractor setting.

A key impediment to providing security is distributional uncertainty.  We consider two settings, first where the fuzzy extractor knows the precise distribution and second the fuzzy extractor knows the distribution comes from some family.  We show that fuzzy min-entropy is a sufficient condition in the first setting and insufficient in the second setting.  That is, there is families of distributions with fuzzy min-entropy that no fuzzy extractor can derive keys from.  Our results are information-theoretic.

\bnote{keep editing}
\end{abstract}

\section{Introduction}
\anote{should our results be seen as justifying the distribution specific approach?}
\anote{make the point that narrowing the family $\mathcal{W}$ makes the security guarantees much more brittle}
Usually, authentication requires some high quality secret.  Traditionally, this secret is assumed to have high min-entropy.  However, many sources with sufficient entropy for authentication present an additional problem: noise~\cite{daugman2004,zviran1993comparison,brostoff2000passfaces,ellison2000protecting,mayrhofer2009shake,monrose2002password,pappu2002physical,tuyls2006puf,gassend2002silicon,suh2007physical,bennett1988privacy}.  That is, when the physical instantiation of the source is read multiple times, readings are close~(according to some metric) but not identical.  To use such sources in an authentication scheme it is often necessary to remove this noise and derive the same key from the initial and subsequent readings.

@@

A fuzzy extractor performs both these tasks non interactively~\cite{DBLP:journals/siamcomp/DodisORS08}.  %Consider some distribution $W$.%The goal of a fuzzy extractor is to derive stable keys from a distribution $W$ whenever the two readings~(denoted $w$ and $w'$ respectively) are within distance $t$.  
A fuzzy extractor consists of two algorithms: generate~($\gen$) which takes an initial reading $w$ and produces $key$ along with a nonsecret helper value $p$.  The reproduce~($\rep$) algorithm takes the subsequent reading $w'$ along with the helper value $p$ to reproduce $key$.  The $key$ should be uniform even if an adversary has observed $p$.  The correctness guarantee is that the key is reproduced if the distance between $w$ and $w'$ is at most $t$.  For any security the initial readings must come from a high min-entropy distribution $W$.

Traditionally, fuzzy extractors use a secure sketch to perform information reconciliation~(mapping $w'$ back to $w$) and a randomness extractor~\cite{nisan1993randomness} to transform $w$ into a uniform key.  The security losses incurred in information-reconciliation dominate for most sources.\footnote{The task of privacy amplification is well-understand.  We have bounds on the length of a key extractible from a source with entropy and extractors that match this bound.}  A secure sketch is a pair of algorithms: $\sketch$ takes $w$ and produces a non secret value $ss$, $\rec$ takes $ss$ and a nearby value $w'$ and outputs the original reading $w$.  
Known constructions of fuzzy extractors and secure sketches incur significant losses.  In particular, for most practical noisy sources, there are no known secure sketches or fuzzy extractors.  In this work, we ask whether this is inherent.  We identify a key problem: uncertainty in the distribution of $W$.

@@

Traditional fuzzy extractors and secure sketches operate in the worst case and incur entropy loss of the size of the ball to be error-corrected.  This loss is usually linear in the input length of the string.  While, this is optimal for the uniform distribution, it is far from optimal for distributions that arise in practice.  As an example, many examples that arise in practice there are more error patterns than starting entropy.  Traditional fuzzy extractors provide no guarantee for sources of this type.  The work of Canetti et al.~\cite{canetti2014key} constructs fuzzy  extractors for some distributions with more errors than entropy.  This leads to the following natural question: \emph{when is key derivation from noisy sources possible?}  To help answer this question, we identify a key problem: uncertainty in the distribution of $W$.

\paragraph{The price of distributional uncertainty}
Fuzzy extractors are designed with limited knowledge of the distributions that will later be used.  Traditionally, the distribution $W$ is assumed to be an arbitrary distribution with min-entropy.  Making minimal assumptions about the distribution $W$ is desirable for two reasons: 1) for high entropy distributions, only a small part of the distribution has been observed 2) the adversary may have some side information about the distribution $W$ and thus the adversaries view of the distribution is different than that of the algorithm designer.  Fuzzy extractors support a family of distributions $\mathcal{W}$ and hope that the target distribution $W\in\mathcal{W}$.

In this work, we show that this uncertainty comes at a significant cost.  We consider two settings, first where the algorithms $\gen, \rep$ know the probability mass function of the distribution $W$.  We call this setting \emph{distribution aware}.  Second, we consider the traditional setting where security holds for a family of distributions $\mathcal{W}$.  To distinguish it from the prior setting, we call it the \emph{distribution-oblivious} setting.
\anote{need to mention early work on trying to do distribution specific. For example, Juels and Wattenberg, Tuyls et al.}

@@

\paragraph{Positive Results} %We first consider the \emph{distribution-aware} setting.  
We show a distribution-aware secure sketch for every source with super-logarithmic fuzzy min-entropy with super-logarithmic output entropy~(\thref{cons:leveling}).  Using the sketch-and-extract construction, there is a distribution-aware fuzzy extractor with super-logarithmic key output length.\footnote{Assuming exponential-hard pseudorandom generators, a super-logarithmic length key can be expanded to an arbitrary length.  In this work, we assume the goal is to produce a super-logarithmic length key.  In the information-theoretic setting, this can be seen as a minimum requirement for security.}  Thus, in the distribution-aware setting, fuzzy min-entropy is both a necessary and sufficient condition for security.

Our constructions are feasibility results and do not run in polynomial time.  In the distribution-aware setting this seems necessary as most distributions have exponential description size.  The power of the distribution-aware setting is that the algorithms can access this description.  It seems unlikely that algorithms with access to a polynomial size part of the distribution description can replicate these results.

\paragraph{Negative Results}
The situation is very different in the distribution-oblivious setting.  We show a family of distributions~(each with super-logarithmic fuzzy min-entropy) such that any secure sketch for the family the remaining entropy is at most $1$~(\thref{thm:imposs sketch}).  This result only requires the secure sketch to correct a constant number of errors and to be correct $75\%$ of the time.

@@

Both of negative results use families of distributions that are well-spread: all points of the distribution lie far apart.  Intuitively, this should be the easiest type of family to perform information-reconciliation.  In the distribution-aware setting, information-reconciliation requires no information.

\paragraph{Computational security?}
Our negative result on information-theoretic secure sketches extends to secure sketches where $w$ has pseudoentropy~\cite{DBLP:journals/siamcomp/HastadILL99} using a result of Fuller et al.~\cite{fuller2013computational}.  Secure sketches that provide computational unpredictability of $w$ are constructible from semantically secure multi-linear encodings~\cite{BitanskyCKP14}.  Semantically secure multi-linear encodings are a very strong assumption and unpredictability is weaker than pseudoentropy.  Converting unpredictability to a key requires a special type of extractor known as a reconstructive randomness extractor~\cite{barak-computational, DBLP:conf/eurocrypt/HsiaoLR07}.

Computational fuzzy extractors offer more hope for avoiding the impossibility result.  Computational fuzzy extractors for all distributions are implied by semantically secure multi-linear encodings~\cite{BitanskyCKP14}.  Furthermore, there are computational fuzzy extractors secure for the specific family of distributions used in the counter-example under weaker assumptions~\cite[Construction 5.3]{canetti2014key}.

\bnote{compare this with universal decoding?}

@@

\subsection{Fuzzy min-entropy: a necessary condition}
\label{sec:minimal conditions}
